### 2019-05-22 1_introduction.tex
New text for GPU comparison and other tweaks.

To address the second issue and isolate the contribution of the programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG). Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
performance (e.g., FLOPS or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency. For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.

>>>>

To address the second issue and isolate the contribution of the programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG). Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
performance (e.g., FLOP/s or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency. For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.

### 2019-05-22 2_metg.tex
Apply Pat's comments.
New text for GPU comparison and other tweaks.

The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
the smallest average task granularity (i.e., task duration) such that $A$
achieves overall efficiency of at least 50\%. For example, in
compute-bound applications efficiency can be measured as the
percentage of the available FLOPS achieved. On Cori with 1.26 TFLOPS available per Haswell node, METG(50\%) corresponds to
the smallest task granularity achieved while maintaining at least 0.63
TFLOPS. Absolute (rather than relative) efficiency is preferred
because it better quantifies the performance tradeoff made by
programming systems that reserve processor cores to implement asynchronous operations (such as data movement).

@@

To calculate METG, the data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOPS achieved) and task
granularity (i.e., $\text{wall time} \times \text{number of
  cores}/\text{number of tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
broadly to be any continuously-executing unit of application code,
and thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
compute-bound kernel included in the Task Bench implementation,
described in more detail in Section~\ref{sec:task-bench}.

In Figure~\ref{fig:efficiency-mpi},
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. However as task
granularity shrinks further, efficiency drops more rapidly, leading
to an asymptote as runtime overhead dominates useful work.

>>>>

The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
the smallest average task granularity (i.e., task duration) such that $A$
achieves overall efficiency of at least 50\%. For example, in
compute-bound applications efficiency can be measured as the
percentage of the available FLOP/s achieved. On Cori with 1.26 TFLOP/s available per Haswell node, METG(50\%) corresponds to
the smallest task granularity achieved while maintaining at least 0.63
TFLOP/s. Absolute (rather than relative) efficiency is preferred
because it better quantifies the performance tradeoff made by
programming systems that reserve processor cores to implement asynchronous operations (such as data movement).

@@

To calculate METG, the data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOP/s achieved) and task
granularity (i.e., $\text{wall time} \times \text{number of
  cores}/\text{number of tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
broadly to be any continuously-executing unit of application code,
and thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
compute-bound kernel included in the Task Bench implementation,
described in more detail in Section~\ref{sec:task-bench}.

In Figure~\ref{fig:efficiency-mpi},
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. However as task
granularity shrinks further, efficiency drops more rapidly, leading
to a vertical asymptote as runtime overhead dominates useful work.

### 2019-05-22 5_evaluation.tex
New text for GPU comparison and other tweaks.

Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOPS of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

The vast majority of systems achieve or nearly achieve peak FLOPS. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOPS compared
to systems which share all cores between the application and runtime. Some of the
slower systems struggle to achieve peak FLOPS, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this configuration ran for over 6 hours.

Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
peak FLOPS) plotted against task granularity. This plot is
used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

\input{e8_weak}
\input{e9_strong}

@@

\input{e5_imbalance}
\input{e6_communication}

@@

% FIXME: TODO

>>>>

Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOP/s of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems which share all cores between the application and runtime. Some of the
slower systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this configuration ran for over 6 hours.

Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
peak FLOP/s) plotted against task granularity. This plot is
used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

\input{e8_weak}
\input{e9_strong}
\input{e5_imbalance}

@@

@@

Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which compares favorably with the officially reported number
$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
FLOP/s. Note that the x-axis is measured in the number of FLOP
performed during the run, as the GPU kernels must perform more work to
achieve peak performance.

Our MPI+CUDA implementation uses an offload model in which input and
output data is copied to and from the GPU on each time step. We tested
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel and thus achieve
some overlap between communication and computation. We see that the
overdecomposed \lstinline{w4} configuration achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data to and from the GPU dominates at small task
granularities. Thus at smaller problem sizes the CPU achieves higher
performance.

### 2019-05-22 e5_imbalance.tex
Tweak spacing around figure captions.
Apply Alex's and George's comments.
New text for GPU comparison and other tweaks.

\caption{Task granularity vs efficiency under load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Lower is better.\label{fig:efficiency-imbalance}}
\end{figure}

>>>>

\vspace{-0.15cm}
\caption{Efficiency vs task granularity under load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Higher is better.\label{fig:efficiency-imbalance}}
\vspace{-0.1cm}
\end{figure}

### 2019-05-22 e6_communication.tex
Apply Alex's and George's comments.
New text for GPU comparison and other tweaks.

\vspace{-0.25cm}
\caption{Task granularity vs efficiency for varying communication (spread pattern, 5 dependencies per task, 4 task graphs, 64 nodes). Lower is better.\label{fig:efficiency-communication}}
\vspace{-0.05cm}
\end{figure*}

>>>>

\vspace{-0.25cm}
\caption{Efficiency vs task granularity for varying communication (spread pattern, 5 dependencies per task, 4 task graphs, 64 nodes). Higher is better.\label{fig:efficiency-communication}}
\vspace{-0.05cm}
\end{figure*}
